On exponential sums arising from the classical doubling method

### 2025-04-02 main.tex
Adding notation for Gauss sums of Tori

% Finite field macros
\newcommand{\FieldNorm}[2]{\mathrm{N}_{#1:#2}}
\newcommand{\aFieldNorm}{\mathrm{N}}
\newcommand{\finiteField}{\mathbb{F}}
\newcommand{\quadraticExtension}{\mathbb{E}}
\newcommand{\finiteFieldExtension}[1]{\finiteField_{#1}}
\newcommand{\quadraticFieldExtension}[1]{\quadraticExtension_{#1}}
\newcommand{\NormOneGroup}[1]{\finiteFieldExtension{#1}^{\aFieldNorm = 1}}
\newcommand{\algebraicClosure}[1]{\overline{#1}}
\newcommand{\Galois}{\operatorname{Gal}}
\newcommand{\Frobenius}{\operatorname{Fr}}
\newcommand{\restrictionOfScalars}[3]{\operatorname{Res}_{#1 \slash #2}{#3}}
\newcommand{\multiplcativeScheme}{\algebraicGroup{G}_m}
\newcommand{\affineLine}{\mathbb{A}^1}
\newcommand{\squareMatrix}{\operatorname{Mat}}
\newcommand{\Mat}[2]{\operatorname{Mat}_{#1 \times #2}}
\newcommand{\GaussSum}[2]{\mathcal{G}\left(#1, #2\right)}
\newcommand{\GaussSumSingleCharacter}[2]{\tau\left(#1, #2\right)}
\newcommand{\dblJacobiSum}[2]{\mathcal{J}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblJacobiSum}[2]{\mathcal{J}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\negDblJacobiSum}[2]{\mathcal{J}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\JacobiKernel}[1]{\Phi_{#1}}
\newcommand{\posJacobiKernel}[1]{\Phi^{+}_{#1}}
\newcommand{\negJacobiKernel}[1]{\Phi^{-}_{#1}}
\newcommand{\genJacobiKernel}[1]{\Phi^{\pm}_{#1}}
\newcommand{\genHermitianJacobiKernel}[2]{\Phi^{\pm}_{#1,#2}}

\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSumScalar}[2]{\mathrm{J}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblJacobiSumScalar}[2]{\mathrm{J}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\negDblJacobiSumScalar}[2]{\mathrm{J}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualJacobiSumScalar}[2]{\mathrm{j}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblPosVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblNegVirtualJacobiSumScalar}[2]{\mathrm{j}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\dblLanglandsGammaFactorSpace}[4]{\gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GaussSumCharacter}[3]{\tau\left(#1 \times #2, #3\right)}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}
\newcommand{\LusztigSeries}[2]{\mathcal{E}\left(#1, (#2)\right)}
\newcommand{\characteristicPolynomial}{\operatorname{CharPoly}}
\newcommand{\DualFrobeniusFixedPoints}[2][\Frobenius^{\ast}]{\algebraicGroup{#2}^{\ast #1}}
\newcommand{\FrobeniusFixedPoints}[2][\Frobenius]{\algebraicGroup{#2}^{#1}}
\newcommand{\CharacterLattice}[1]{X^{\ast}\left(#1\right)}
\newcommand{\CocharacterLattice}[1]{X_{\ast}\left(#1\right)}
\newcommand{\SymmetricGroup}{\mathfrak{S}}
\newcommand{\Kernel}{\operatorname{Ker}}

@@

We are now ready to state Kondo's result.
%test
\begin{theorem}[Kondo {\cite{Kondo1963}}]\label{thm:KondoClassical}
	Suppose that $\tau$ is an irreducible representation of $\GL_n\left(\finiteField\right)$ with cuspidal support $\left\{\tau_1,\dots, \tau_r\right\}$, where $n_1 + \dots + n_r = n$, and where for every $j$, $\tau_j$ is an irreducible cuspidal representation of $\GL_{n_j}\left(\finiteField\right)$ corresponding to the Frobenius orbit of a regular character $\theta_j \colon \multiplicativegroup{\finiteFieldExtension{n_j}} \to \multiplicativegroup{\cComplex}$. Then
	$$\GaussSumScalar{\tau \times \chi}{\fieldCharacter} = \left(-1\right)^n \cdot \prod_{j=1}^r \GaussSumCharacter{\theta_j}{\chi}{\fieldCharacter_{n_j}}.$$
\end{theorem}

@@

\begin{theorem}\label{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}
	Suppose that $\algebraicGroup{T}_{\algebraicGroup{H}}$ is a maximal rational torus of $\algebraicGroup{H}$ that corresponds to the pair of partitions $\left(\lambda^+, \lambda^-\right)$ and let $\algebraicGroup{T}_{\algebraicGroup{G}} \coloneq \algebraicGroup{G} \cap \algebraicGroup{T}_{\algebraicGroup{H}}$ and $T = T_G = \algebraicGroup{T}_{\algebraicGroup{G}}^{\Frobenius}$. Suppose that $\theta_{H} \colon \algebraicGroup{T}_{\algebraicGroup{H}}^{\Frobenius} \to \multiplicativegroup{\cComplex}$ is a character and let $$\theta = \theta_{\algebraicGroup{G}} \coloneq \theta_{\algebraicGroup{H}} \restriction_{T_G} \colon T_G \to \multiplicativegroup{\cComplex}.$$ Then
	\begin{align*}
		 \dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = R_{T,\theta}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) g_{T'}\left(\chi, \theta', \fieldCharacter\right),
	\end{align*}
	and \begin{align*}
		\dblNegVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \begin{dcases}
			0 & \dim_{\finiteField} \hermitianSpace \text{ is odd.}\\
			\dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} \cdot \theta\left(-1\right) & \text{ otherwise.}
		\end{dcases}
	\end{align*}
\end{theorem}
\begin{proof}
	Notice that if $\dim_{\finiteField} \hermitianSpace$ is odd then $\negJacobiKernel{\chi}\left(t\right) = 0$ for any $t \in T_{G}$, as $\idmap_{\hermitianSpace} - t$ always has determinant zero.

>>>>

% Finite field macros
\newcommand{\FieldNorm}[2]{\mathrm{N}_{#1:#2}}
\newcommand{\aFieldNorm}{\mathrm{N}}
\newcommand{\finiteField}{\mathbb{F}}
\newcommand{\quadraticExtension}{\mathbb{E}}
\newcommand{\finiteFieldExtension}[1]{\finiteField_{#1}}
\newcommand{\quadraticFieldExtension}[1]{\quadraticExtension_{#1}}
\newcommand{\NormOneGroup}[1]{\finiteFieldExtension{#1}^{\aFieldNorm = 1}}
\newcommand{\algebraicClosure}[1]{\overline{#1}}
\newcommand{\Galois}{\operatorname{Gal}}
\newcommand{\Frobenius}{\operatorname{Fr}}
\newcommand{\restrictionOfScalars}[3]{\operatorname{Res}_{#1 \slash #2}{#3}}
\newcommand{\multiplcativeScheme}{\algebraicGroup{G}_m}
\newcommand{\affineLine}{\mathbb{A}^1}
\newcommand{\squareMatrix}{\operatorname{Mat}}
\newcommand{\Mat}[2]{\operatorname{Mat}_{#1 \times #2}}
\newcommand{\GaussSum}[2]{\mathcal{G}\left(#1, #2\right)}
\newcommand{\GaussSumSingleCharacter}[2]{\tau\left(#1, #2\right)}
\newcommand{\GaussSumSingleTorusCharacter}[3]{\tau_{#1}\left(#2, #3\right)}
\newcommand{\dblJacobiSum}[2]{\mathcal{J}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblJacobiSum}[2]{\mathcal{J}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\negDblJacobiSum}[2]{\mathcal{J}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\JacobiKernel}[1]{\Phi_{#1}}
\newcommand{\posJacobiKernel}[1]{\Phi^{+}_{#1}}
\newcommand{\negJacobiKernel}[1]{\Phi^{-}_{#1}}
\newcommand{\genJacobiKernel}[1]{\Phi^{\pm}_{#1}}
\newcommand{\genHermitianJacobiKernel}[2]{\Phi^{\pm}_{#1,#2}}

\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSumScalar}[2]{\mathrm{J}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblJacobiSumScalar}[2]{\mathrm{J}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\negDblJacobiSumScalar}[2]{\mathrm{J}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualJacobiSumScalar}[2]{\mathrm{j}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblPosVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblNegVirtualJacobiSumScalar}[2]{\mathrm{j}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\dblLanglandsGammaFactorSpace}[4]{\gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GaussSumCharacter}[3]{\tau\left(#1 \times #2, #3\right)}
\newcommand{\GaussSumTorusCharacter}[4]{\tau_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}
\newcommand{\LusztigSeries}[2]{\mathcal{E}\left(#1, (#2)\right)}
\newcommand{\characteristicPolynomial}{\operatorname{CharPoly}}
\newcommand{\DualFrobeniusFixedPoints}[2][\Frobenius^{\ast}]{\algebraicGroup{#2}^{\ast #1}}
\newcommand{\FrobeniusFixedPoints}[2][\Frobenius]{\algebraicGroup{#2}^{#1}}
\newcommand{\CharacterLattice}[1]{X^{\ast}\left(#1\right)}
\newcommand{\CocharacterLattice}[1]{X_{\ast}\left(#1\right)}
\newcommand{\SymmetricGroup}{\mathfrak{S}}
\newcommand{\Kernel}{\operatorname{Ker}}

@@

Given a torus $T \subset \GL_n\left(\finiteField\right)$ of the form $T \cong \multiplicativegroup{\finiteFieldExtension{\lambda}}$ where $\lambda \vdash n$ is a partition and given characters $\alpha \colon \multiplicativegroup{\finiteFieldExtension{\lambda}} \to \multiplicativegroup{\cComplex}$ and $\chi \colon T \to \multiplicativegroup{\cComplex}$, we define $$\GaussSumTorusCharacter{T}{\alpha}{\chi}{\fieldCharacter} = \left(-1\right)^{\lengthof\left(\lambda\right)} q^{-\frac{n}{2}} \sum_{t \in T} \alpha^{-1}\left(t\right) \chi^{-1}\left(\det t\right) \fieldCharacter\left(\trace t\right).$$
Notice that if $\alpha = \alpha_1 \times \dots \times \alpha_{\lengthof\left(\lambda\right)}$, where $\alpha_j \colon \multiplicativegroup{\finiteFieldExtension{\lambda_j}} \to \multiplicativegroup{\cComplex}$ is a character for every $j$, then
$$\GaussSumTorusCharacter{T}{\alpha}{\chi}{\fieldCharacter} = \prod_{j=1}^{\lengthof\left(\lambda\right)} \GaussSumCharacter{\alpha}{\chi}{\fieldCharacter_{\lambda_j}}.$$

We are now ready to state Kondo's result.
\begin{theorem}[Kondo {\cite{Kondo1963}}]\label{thm:KondoClassical}
	Suppose that $\tau$ is an irreducible representation of $\GL_n\left(\finiteField\right)$ with cuspidal support $\left\{\tau_1,\dots, \tau_r\right\}$, where $n_1 + \dots + n_r = n$, and where for every $j$, $\tau_j$ is an irreducible cuspidal representation of $\GL_{n_j}\left(\finiteField\right)$ corresponding to the Frobenius orbit of a regular character $\theta_j \colon \multiplicativegroup{\finiteFieldExtension{n_j}} \to \multiplicativegroup{\cComplex}$. Then
	$$\GaussSumScalar{\tau \times \chi}{\fieldCharacter} = \left(-1\right)^n \cdot \prod_{j=1}^r \GaussSumCharacter{\theta_j}{\chi}{\fieldCharacter_{n_j}}.$$

	If $\tau$ is an irreducible representation such that $\innerproduct{\trace \tau}{R_{T, \alpha}} \ne 0$ for some maximal torus $T$ of $\GL_n\left(\finiteField\right)$ and some character $\alpha \colon T \to \multiplicativegroup{\cComplex}$ then $$\GaussSumScalar{\tau \times \chi}{\fieldCharacter} = \left(-1\right)^n \GaussSumTorusCharacter{T}{\alpha}{\chi}{\fieldCharacter}.$$
\end{theorem}

@@

\begin{theorem}\label{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}
	Suppose that $\algebraicGroup{T}_{\algebraicGroup{H}}$ is a maximal rational torus of $\algebraicGroup{H}$ that corresponds to the pair of partitions $\left(\lambda^+, \lambda^-\right)$ and let $\algebraicGroup{T}_{\algebraicGroup{G}} \coloneq \algebraicGroup{G} \cap \algebraicGroup{T}_{\algebraicGroup{H}}$ and $T = T_G = \algebraicGroup{T}_{\algebraicGroup{G}}^{\Frobenius}$. Suppose that $\theta_{H} \colon \algebraicGroup{T}_{\algebraicGroup{H}}^{\Frobenius} \to \multiplicativegroup{\cComplex}$ is a character and let $$\theta = \theta_{\algebraicGroup{G}} \coloneq \theta_{\algebraicGroup{H}} \restriction_{T_G} \colon T_G \to \multiplicativegroup{\cComplex}.$$ Then
	\begin{align*}
		 \dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = R_{T,\theta}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) \GaussSumTorusCharacter{T'}{\theta'}{\chi}{\fieldCharacter},
	\end{align*}
	and \begin{align*}
		\dblNegVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \begin{dcases}
			0 & \dim_{\finiteField} \hermitianSpace \text{ is odd,}\\
			\dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} \cdot \theta\left(-1\right) & \text{ otherwise.}
		\end{dcases}
	\end{align*}
\end{theorem}
\begin{proof}
	Notice that if $\dim_{\finiteField} \hermitianSpace$ is odd then $\negJacobiKernel{\chi}\left(t\right) = 0$ for any $t \in T_{G}$, as $\idmap_{\hermitianSpace} - t$ always has determinant zero.
